\class : Problem Set \psetnum

### 2022-05-16 531_stars/pset4/solutions.tex
finished HW4 stars

\question{Part c - Density of bi-stability disk}
\answer{
    A rotation-induced bi-stability disk must by definition have a lower effective temperature than the wind in the poles and it also has a lower terminal wind velocity. The lower temperature naturally leads to higher densities and the lower velocity allows more gas to remain in the disk.
    \todo{check this part, particularly the velocity bit}
}

@@

    \textit{Higher mass limit:} The mass needed to create a BH would need to increase due to the increased mass loss due to rotation. This means that less mass is available to go into the creation of a remnant and so stars that would have resulted in BHs instead result in NSs.

    \textit{Lower mass limit:} However in addition to the overall mass loss, the mass of the helium core instead increases for rapidly-rotating stars due to the induced mixing. A larger core mass can also lead to the larger remnant mass and thus more easily create a BH with a given initial mass.

    Overall, I think that\dots
    \todo{Check with Emily what's more important}
}

>>>>

\question{Part c - Density of bi-stability disk}
\answer{
    The density of the gas scales as
    \begin{equation}
        \rho \sim \frac{\dot{M}}{v_{\infty}}
    \end{equation}
    For a rotation-induced bi-stability disk, the mass loss is higher than at the poles at its slightly lower $T_{\rm eff}$ places it just below a bi-stability jump (see Section 15.3). Moreover, the terminal wind velocity ($v_\infty$) is higher on the lower side of a bi-stability jump (see Section 15.2.2). The combination of these effects leads to a significant increase in density in the disk relative to the poles.
}

@@

    \begin{itemize}
        \item \textit{Case for higher mass limit:} The mass needed to create a BH would need to increase to compensate for the increased mass loss due to rotation. Otherwise, less mass is available to go into the creation of a remnant and so stars that would have resulted in BHs instead result in NSs.
        \item \textit{Case for lower mass limit:} Though the overall mass is decreased for a rapidly-rotating star, the mass of the helium core \textit{increases} due to the induced mixing. A larger core mass leads to a larger remnant mass and thus more easily create a BH with a given initial mass.
    \end{itemize}
    Overall, the winner here is the latter reasoning. Core mass is the most important parameter for determining the destiny of any massive star and so if all stars were rapidly-rotating the mass limit between NSs and BHs would decrease.
}
